---
title: "Using Color"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=F, message=F}
library(dplyr)
library(ggethos)
library(ggplot2)
```

In this article we will revisit the basic usage adding color.

## Outside `aes()`

You are free to decide how to color the segments in the ethograms when using `geom_ethogram()`.

```{r color-salmon}
ggplot(wombats, aes(y = wombat, 
                     behaviour = behaviour)) +
     geom_ethogram(remove_nas = FALSE, color="salmon")
```


## Inside `aes()`

We can also choose to use color based on any column present in our data. For example, we can color by `wombat`:

```{r color-wombat}
ggplot(wombats, aes(y = wombat, 
                     behaviour = behaviour)) +
     geom_ethogram(aes(color=wombat), 
                   remove_nas = FALSE)
```

Perhaps the most common use would be to color using the `behaviour`:

```{r color-behaviour}
ggplot(wombats, aes(y = wombat, 
                     behaviour = behaviour)) +
     geom_ethogram(aes(color=behaviour), 
                   remove_nas = FALSE)
```

Note that we have been using `remove_nas=FALSE`, while we would produce a different output when using the default:

```{r color-behaviour-no-nas}
ggplot(wombats, aes(y = wombat, 
                     behaviour = behaviour)) +
     geom_ethogram(aes(color=behaviour))
```

## Summary

In this article we will revisit the basic usage adding color. To explore further uses see [Using Trials](articles/trials.html)

## Session Info

```{r session-info}


```

