---
title: "Using Trials"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=F, message=F}
library(dplyr)
library(ggethos)
library(ggplot2)
```

In many situations, behaviour is collected in trials. For this reason, we have provided `trials` in `wombats` and the `align_trials` parameter in `geom_ethogram()`.

## Understanding Trials

Let's explore the nature of the dataset. This is how it looks like when not visualizing trials.

```{r seconds-no-trials}
ggplot(wombats, 
       aes(x=seconds, 
           y = wombat, 
           behaviour = behaviour,
           color=behaviour)) +
  geom_ethogram() 
```

Here's when we facet by trial:

```{r seconds-facet-trials}
ggplot(wombats, 
       aes(x=seconds, 
           y = wombat, 
           behaviour = behaviour,
           color = behaviour)) +
  geom_ethogram() +
  facet_wrap(~trial)
```

We can now see the issue: the trials were collected with a continuous clock and we would like to align them to common time (i.e., have all trials start at zero). This is when `align_trials` comes handy:

```{r seconds-facet-trials-aligned}
ggplot(wombats, 
       aes(x=seconds, 
           y = wombat, 
           behaviour = behaviour,
           color = behaviour)) +
  geom_ethogram(align_trials=TRUE) +
  facet_wrap(~trial)
```

Here's an another example. Let's say we wanted to facet by `wombat` and have `trial` be on the `y` axis instead.

```{r trial-y}
ggplot(wombats, 
       aes(x=seconds, 
           y = trial, 
           behaviour = behaviour,
           color = behaviour)) +
    geom_ethogram() +
    facet_wrap(~wombat)

```
We can easily align trials again using `align_trials=T`:

```{r trial-y-aligned}
ggplot(wombats, 
       aes(x=seconds, 
           y = trial, 
           behaviour = behaviour,
           color = behaviour)) +
    geom_ethogram(align_trials = T) +
    facet_wrap(~wombat)

```

## Summary 

In this article, we have covered basic functionality of aligning trials by using the `align_trials` argument.

## Session Info

```{r}
sessioninfo::session_info()
```

